---
title: "Algorithms assessment via cluster"
author: "Shun Bi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Algorithms assessment via cluster}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

This vignette mainly shows how to use function `Assessment_via_cluster()` for comparing several algorithms if membership values (returned by `FCM.new()`) are given. I am also happy to introduce the function `run_all_Chla_algorithms()` for running all supported Chla estimation algorithms (run `Chla_algorithms_name()` to see the algorithm names). 

Here we use the `Nechad2015` dataset as the example. Similarly, 100 samples are randomly selected from the original set to save the running time. You can test the whole set if you want. 

# Preparing the data set

Based on some prior knowledge (shown in other vignettes), we set the cluster number (i.e., `nb`) as `4` and run the `FCMm` with normalized spectra.

```{r fig.height=3, fig.width=4, message=FALSE, warning=FALSE}
library(FCMm) 
library(ggplot2) 
library(magrittr)
library(stringr)
data("Nechad2015")
x <- Nechad2015[,3:11]
wv <- gsub("X","",names(x)) %>% as.numeric
set.seed(1234) # Set this seed so that you can re-produce them
w <- sample(1:nrow(x), 150) # sample some rows of the orignal data, you can test the whole set if you like
x <- x[w, ]
names(x) <- wv
nb = 4 # Obtained from the vignette "Cluster a new dataset by FCMm"
set.seed(1234)
FD <- FuzzifierDetermination(x, wv, do.stand=FALSE)
result <- FCM.new(FD, nb, fast.mode = TRUE)
p.spec <- plot_spec(result, show.stand=TRUE)
print(p.spec$p.cluster.spec)
Chla <- Nechad2015$X.Chl_a..ug.L.[w]
Chla[Chla >= 999] <- NA
```

# Algorithm assessment via cluster

`BR_Gil10`, `OC4_OLCI`, `OCI_Hu12`, and `NDCI_Mi12` are selected for estimating the Chla concentration. Since there are several `NA` values in the selected measured Chla concentration (also termed as actual value) from `Nechad2015`, observations including `NA` values are removed from the input set.

## Chlorophyll a concentration estimation
```{r}
dt_Chla <- run_all_Chla_algorithms(x) %>% as.data.frame
dt_Chla <- data.frame(
  Chla_true = Chla, 
  BR_Gil10 = dt_Chla$BR_Gil10,
  OC4_OLCI = dt_Chla$OC4_OLCI,
  OCI_Hu12 = dt_Chla$OCI_Hu12,
  NDCI_Mi12= dt_Chla$NDCI_Mi12
) %>% round(3)
w = which(!is.na(dt_Chla$Chla_true))
dt_Chla = dt_Chla[w,]
memb = result$res.FCM$u[w,] %>% round(4)
```

## Scatter plot for the selected algorithms

Shown in the scatter plots, the `OC4_OLCI` and `OCI_Hu12` performed better than other two algorithms across all water clusters.

```{r fig.height=3, fig.width=4}
data.frame(dt_Chla, cluster = result$res.FCM$cluster[w] %>% as.character) %>%
  reshape2::melt(., id=c("Chla_true","cluster")) %>% 
  ggplot() + 
  geom_point(aes(Chla_true, value, color=cluster, group=variable)) + 
  geom_abline(slope=1, intercept=0, linetype=2) + 
  facet_wrap(~variable) + 
  scale_x_log10(limits=c(0.1, 200), name="Actual value") + 
  scale_y_log10(limits=c(0.1, 200), name="Predicted value") + 
  scale_color_brewer(palette = "Set1", name="Cluster") + 
  theme_bw() + 
  theme(text = element_text(size=13))
```

## Algorithms assessment via cluster in soft and hard mode

There is no need `cluster` information in the function `Assessment_via_cluster()` because the cluster will be found via membership matrix (i.e., the parameter `memb`). I fully recommend to set `na.process = TRUE` so that the function will count the valid result. `plot.col` is very useful when you want to visually see the performance comparison between different algorithms.

```{r fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
colnames(memb) <- 1:4
Asses_soft <- Assessment_via_cluster(pred = dt_Chla[,-1],
                                     meas = dt_Chla[,1],
                                     memb = memb,
                                     log10 = TRUE,
                                     hard.mode = FALSE,
                                     na.process = TRUE,
                                     plot.col = TRUE)

Asses_hard <- Assessment_via_cluster(pred = dt_Chla[,-1],
                                     meas = dt_Chla[,1],
                                     memb = memb,
                                     log10 = TRUE,
                                     hard.mode = TRUE,
                                     na.process = TRUE,
                                     plot.col = TRUE)

Asses_soft$res_plot_facet
knitr::kable(Asses_soft$MAE %>% round(3))
knitr::kable(Asses_soft$MAPE %>% round(2))

Asses_hard$res_plot_facet
knitr::kable(Asses_hard$MAE %>% round(3))
knitr::kable(Asses_hard$MAPE %>% round(2))
```

