---
title: "How to apply the default FCMm to new data?"
author: "Shun Bi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to apply the default FCMm to new data?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	collapse = TRUE,
	comment = "#>"
)
```

# Lists of vignettes showing how to use FCMm

Dear users, the four vignettes of `FCMm` were provided in this package to demonstrate the usage and application to obtain the cluster number, to train the user-defined data set, to apply the default FCMm to new data (such as Chla concentration estimation), and the imagery data.

Here are the step-by-step demos:

 - [Demo_1: How to determine the cluster number?](./doc/Cluster_number_determination.html)
 - [Demo_2: How to train the fuzzy cluster?](./doc/Fuzzy_cluster_training.html)
 - [**Demo_3: How to apply the default FCMm to new data?**](./doc/New_Data_Running_FCMm.html)
 - [Demo_4: The application of FCMm to imagery data.](./doc/Raster_Running_FCMm.html)

# Preparing the data set

This demo shows how to apply `FCMm` to a new data set if users just want to run it by the default cluster spectra and fuzzifier. The required data sets are `WaterSpec35` and `Bi_clusters`. The former is a dataset with water spectra of inland waters and corresponding Chla concentration (ug/L unit). The latter one is the seven water cluster spectra which trained by FCM-m method (see more details in Bi *et al.* (2019))

**Note_1**: Have to know that `WaterSpec35` is **ONLY** for the testing of `FCMm` in this package. Therefore, if you want to use this dataset to your research, you should contact us to ask for the permit at first via bishun1994@foxmail or liyunmei@njnu.edu.cn. I hope you can understand.

**Note_2**: we also need package `tidyverse` to string and data processing.

**Note_3**: The section `FCM running` used the default Rrs clusters. However, the input of Rrs data may not match the default wavelength settings such as quality control or atmospheric correction failure or so. Anyway, I have provided another section `One more thing` if your wavelength is not matching with the default. I hope you have a good day.

## Dataframe process

```{r message=FALSE, warning=FALSE}
rm(list=ls())
library(FCMm)
library(ggplot2)
library(magrittr)
data("WaterSpec35")
data("Bi_clusters")
Rrs <- WaterSpec35[,3:17]
```

## Data set plots

At first, it is necessary to check out the dataframe situation before clustering (I suppose it is also important to do that in any other researches). The density plots show that the corresponding Chla concentration is severely skewed so it allows us to do a log10-transformed when estimate or validate the candidate algorithms.

Function `plot_spec_from_df` supports a quick look about the spectral dataframe. Just type `help(plot_spec_from_df)` to see more details.

**Note**: The input of `plot_spec_from_df` should be a matrix or data.frame with colnames that could be transformed into the numeric --- as the x-axis of the plot. Since the return of `plot_spec_from_df` is a `ggplot` list, you could modify or add it for your purpose (such as add labs or themes).

```{r fig.height=4, fig.width=6}
qplot(data=WaterSpec35, x=Chla, geom='density', xlab='The raw Chla') + 
  theme_bw() + theme(text = element_text(size=13))
qplot(data=WaterSpec35, x=Chla, geom='density', log='x', xlab='Log10 transformed Chla') + 
  theme_bw() + theme(text = element_text(size=13))
p.spec <- plot_spec_from_df(Rrs) + 
  labs(x='Wavelength (nm)',y=expression(Rrs~(sr^-1))) + 
  theme_bw() + 
  theme(legend.position='none', text=element_text(size=13))
print(p.spec)
```

# FCM running

## Clustering the new data

Function `apply_FCM_m` provides many parameters for users to decide the process for their desired results. In the following case, `result <- apply_FCM_m(Rrs=Rrs)` is equal to `apply_FCM_m(Rrs=Rrs, wavelength=wavelength.default, Rrs_clusters=Rrs_clusters.default, stand=FALSE, default.cluster=TRUE, m_used=1.36, option.plot=TRUE)`. You can see more details by typing `help(apply_FCM_m)`

```{r message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
result <- apply_FCM_m(Rrs=Rrs, option.plot=TRUE)
summary(result)
result$p.group
```

`result` list contains several result by `apply_FCM_M` (partly same to `FD` list):

 - *x*: the raw input Rrs dataframe with unit sr^-1
 - *x.stand*: the standardized Rrs dataframe, if `stand=F`
 - *d*: distance to each cluster
 - *u*: membership values
 - *cluster*: defined by the maximum of membership
 - *quality*: the quality of the cluster results. See more details in `help(apply_FCM_m)`
 - *m.used*: the used value of fuzzifier(m)
 - *K*: cluster number
 - *p.group*: a ggplot list for plotting the cluster result
 - *p.group.facet*: a ggplot list by facet for plotting the cluster result
 - *dt.plot*: dataframe used for ggplot

## Estimating Chla concentration by FCMm

Our package provides a new framework to estimate Chla concentration. The main point is **algorithm blending** which uses the advantages of candidates algorithm for their optimizing water type (such as TBA for turbid water and BR for relatively clear water). The three build-in algorithms used in `FCMm` are `TBA`, `BR`, and `Bloom`. Note that the `Bloom` model was designed for water bloom (or called harmful algal bloom waters with extremely high Chla concentration or scums). Please see more details in `help(FCM_m_Chla_estiamtion)`.

```{r fig.height=4, fig.width=6}
library(reshape2)
dt_Chla <- FCM_m_Chla_estimation(Rrs=data.frame(Rrs665=Rrs$`665`,
                                                Rrs709=Rrs$`708.75`,
                                                Rrs754=Rrs$`753.75`),
                                 U=result$u)
dt_Chla$cluster <- result$cluster %>% as.character
dt_Chla$Chla_true <- WaterSpec35$Chla

options(scipen=10000)

subset(dt_Chla, select=c('cluster','Chla_true','BR','TBA','Bloom','conc.Blend')) %>%
  melt(., id=c('cluster','Chla_true')) %>%
  ggplot(data=.) + 
  geom_point(aes(x=Chla_true,y=value,group=cluster,color=cluster),
             alpha=0.8, size=4) +
  scale_x_log10(limits=c(1,800)) + 
  scale_y_log10(limits=c(1,800)) +
  scale_color_manual(values=heatmaply::RdYlBu(result$K)) + 
  labs(x='True value of Chla concentration (ug/L)',
       y='Estimated value of Chla concentration (ug/L)',
       color='Cluster') + 
  geom_abline(intercept=0, slope=1, linetype=2) + 
  facet_wrap(~variable, nrow=2) + 
  theme_bw() + 
  theme(axis.text.x.bottom = element_text(hjust=1),
        strip.background = element_rect(color="white", fill="white"))

MAPEs <- NULL
i <- 1
for(model in c('BR','TBA','Bloom','conc.Blend')){
  MAPEs[i] <- cal.metrics(x=dt_Chla$Chla_true %>% log10,
                          y=dt_Chla[,model]  %>% log10,
                          name='MAPE',log10=TRUE)
  names(MAPEs)[i] <- model
  i <- i + 1
}
names(MAPEs)[i-1] = 'Blend'
print(MAPEs %>% round(2))

MAEs <- NULL
i <- 1
for(model in c('BR','TBA','Bloom','conc.Blend')){
  MAEs[i] <- cal.metrics(x=dt_Chla$Chla_true,
                         y=dt_Chla[,model],
                         name='MAE',log10=TRUE)
  names(MAEs)[i] <- model
  i <- i + 1
}
names(MAEs)[i-1] = 'Blend'
print(MAEs %>% round(3))
```

It is glad to see the Blending result performs better than others with its log-transformed MAPE = `r MAPEs['conc.Blend'] %>% round(2)` % and MAE = `r MAEs['conc.Blend'] %>% round(2)` ug/L.

# One more thing

As mentioned before, the input Rrs may do not match the default wavelength from `wavelength.default`. This section will show you how to do that kind of clustering by `FCMm`.

Here we suppose the selected bands are `412.5`, `442.5`, `490`, `510`, `560`, `620`, `665`, `673.75`, `708.75`, `753.75`, `865`, and `885`. Note that the `m.used` of `FD` list may change a little if different band number is used. But we assume the difference is negligible (i.e., still equal to the former one `r result$m_used`).

```{r fig.height=4, fig.width=6}
Rrs_sub <- subset(Rrs, select=c(`412.5`,`442.5`,`490`,`510`,
                                `560`,`620`,`665`,`673.75`,
                                `708.75`,`753.75`,`865`,`885`))
wavelength.sub <- c(412.5,442.5,490,510,
                    560,620,665,673.75,
                    708.75,753.75,865,885)
Rrs_clusters.sub <- which(names(Rrs_clusters.default) != 'X400' &
                            names(Rrs_clusters.default) != 'X681' & 
                            names(Rrs_clusters.default) != 'X779') %>%
  Rrs_clusters.default[,.]

# Note the parameter settings in this function `default.cluster=F`
result_sub <- apply_FCM_m(Rrs=Rrs_sub, wavelength=wavelength.sub,
                          Rrs_clusters=Rrs_clusters.sub,
                          stand=F, default.cluster=F, option.plot=T)
result_sub$p.group
```

